---
title: "Aides Drug"
author: "Louie A"
date: "6/15/2023"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

```{r}
Charlie <- filter(triUmp, DrugUser== "yes")
```

```{r eval=FALSE}
Charlie <- Charlie %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->


```{r}
barchartGC(~infected + treatment,data=Charlie)
barchartGC(~infected + treatment,data=Charlie, type="percent")

```
the two graphs show who got the disease 





```{r}
library(ggplot2)

# Assuming "infected" and "treatment" are column names in the data frame "Charlie"

# Bar chart for "infected" variable
ggplot(Charlie, aes(x = infected, fill = treatment)) +
  geom_bar(position = "dodge", color = "black") +
  labs(x = "Infected", y = "Count") +
  ggtitle("Bar Chart of Infected and Treatment") +
  scale_fill_manual(values = c("drug" = "blue", "placebo" = "red")) +
  theme(legend.position = "bottom")

# Bar chart for "treatment" variable
ggplot(Charlie, aes(x = treatment, fill = treatment)) +
  geom_bar(position = "dodge", color = "black") +
  labs(x = "Treatment", y = "Count") +
  ggtitle("Bar Chart of Treatment") +
  scale_fill_manual(values = c("drug" = "blue", "placebo" = "red")) +
  theme(legend.position = "bottom")
```

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table2 <- xtabs(~infected + treatment, data=Charlie)
rowPerc(table2)
colPerc(table2)
```
Those who got aid where at 25% whith the drug taken. Meanwhile 75% took the placebo. Overall those who took the drug or the placebo didnt get infected.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisq.test(table2)
chisqtestGC(table2)
```
So there is a big difference in the p value. Those who took the drug in which 25%, prevented them from getting sick. Those who took the placebo, 75% them got sick.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
today we looked at results of a drug trying to prevent a decease with a placebo effect in place. 
We Also used chat gpt to help out with our data finding but up to a point was a loss but with a help tab in posit.cloud we were able to continue. we concluded that those who took the drug were less likely to get infected than those who took the placebo.